---
title: "MetaVolcanoR: gene and transcript-level meta-analysis"
author: "Izabela Mamede, Bárbara Marinho, Cesar Prada, Carlos Mata-Machado, Helder Nakaya"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{MetaVolcanoR: gene and transcript-level meta-analysis}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r style, echo=FALSE, results="asis", message=FALSE}
knitr::opts_chunk$set(
  tidy    = FALSE,
  warning = FALSE,
  message = FALSE,
  cache   = TRUE
)
```

> 📄 See also: [Protein-level meta-analysis — CPTAC pan-cancer example](Proteomics.html)

# Introduction: Why Meta-Analysis?

Imagine you're studying a disease and find 5 different studies on Gene Expression Omnibus, each identifying
~500 differentially expressed genes (DEGs). But only 50 genes overlap! Which genes are truly important?

**MetaVolcanoR solves this by:**

- Combining evidence across studies
- Identifying consistently perturbed genes
- Visualizing meta-analysis results intuitively

Comparing the expression of genes under a given condition against a reference biological state is usually
applied to identify sets of differentially expressed genes (DEG). These DEG point out the genomic regions
functionally relevant under the biological condition of interest.

Although individual genome-wide expression studies have small signal/noise ratio, today's genomic data
availability usually allows combining differential gene expression results from dozens of independent studies
to overcome this limitation.

Databases such as [GEO](https://www.ncbi.nlm.nih.gov/geo/), [SRA](https://www.ncbi.nlm.nih.gov/sra),
[ArrayExpress](https://www.ebi.ac.uk/arrayexpress/), and [ENA](https://www.ebi.ac.uk/ena) offer systematic
access to vast amounts of transcriptome data. There exists more than one gene expression study for many
biological conditions. This redundancy can be exploited by meta-analysis approaches to reveal genes that
are consistently and differentially expressed under given conditions.

MetaVolcanoR was designed to identify the genes whose expression is consistently perturbed across several
DE tables, and — as this vignette also shows — it applies just as well to other feature types (transcripts,
methylation probes) that share the same identifier / log-fold-change / p-value / confidence-interval
structure.

## Overview

The MetaVolcanoR R package combines differential expression results across studies. It implements three
strategies to summarize differential activity from different studies:

1. **Random Effects Model (REM)** — weights each study by the inverse of its variance and estimates a
   summary fold-change and summary p-value per feature.
2. **A p-value combining approach** — combines per-study p-values with Fisher's method and summarizes the
   fold-change by mean or median.
3. **A vote-counting approach** — counts, per feature, how many studies call it differentially expressed and
   in which direction.

In all three, MetaVolcanoR exploits the volcano-plot reasoning to visualize the meta-analysis result.

## Load library

```{r packages}
library(MetaVolcanoR)
library(dplyr)
library(ggplot2)
```

# Preparing Your Data

MetaVolcanoR requires differential expression results with:

1. **Feature identifiers** (gene names, transcript IDs, probe IDs, ...)
2. **Log2 fold changes**
3. **P-values**
4. **Confidence intervals OR variance** (for the REM method only)

Not sure where to start? Pick the tab below that matches what you already have or which tool produced
your results — each one shows only the code relevant to that path.

## What do you have? {.tabset .tabset-pills}

### CI ready

If your DE results already include confidence interval columns (e.g. `CI.L` and `CI.R`), you're ready to
go — the package will automatically calculate variance from them:

```{r ci_ready, eval=FALSE}
# Your data has CI.L and CI.R columns
meta_rem <- rem_mv(
  diffexp = your_data_list,
  llcol   = "CI.L",   # Left limit of CI
  rlcol   = "CI.R",   # Right limit of CI
  cvar    = TRUE      # Calculate variance from CI (default)
)
```

### Standard error (SE)

Most tools (DESeq2, limma, edgeR) output standard error, not confidence intervals. Convert SE to a 95% CI:

```{r se_to_ci, eval=FALSE}
# From DESeq2 results
deseq_results <- results(dds)  # or after shrinkage
deseq_results$CI.L <- deseq_results$log2FoldChange - 1.96 * deseq_results$lfcSE
deseq_results$CI.R <- deseq_results$log2FoldChange + 1.96 * deseq_results$lfcSE

# From limma results (if you have an SE column)
limma_results$CI.L <- limma_results$logFC - 1.96 * limma_results$SE
limma_results$CI.R <- limma_results$logFC + 1.96 * limma_results$SE
```

### Variance directly

If you have variance (or can calculate it from SE: `variance = SE^2`), use the `vcol` parameter instead:

```{r variance_direct, eval=FALSE}
your_data$variance <- your_data$SE^2

meta_rem <- rem_mv(
  diffexp = your_data_list,
  vcol    = "variance",  # Column name with variance
  cvar    = FALSE        # Don't calculate from CI
)
```

### Test statistic (z or t)

If your DE results have no confidence interval columns, variance can be approximated from the fold-change
and the test statistic (e.g. z-score or moderated t-statistic), using `stat ≈ log2FC / SE`, which gives
`var = (log2FC / stat)^2`:

```{r variance_estimation, eval=FALSE}
diffexp_list <- lapply(diffexp_list, function(df) {
  df %>%
    mutate(var = ifelse(stat == 0 | is.na(stat),
                        NA_real_,
                        (log2FC / stat)^2))
})
```

The `ifelse` guard matters: transcripts where `stat = 0` would otherwise produce infinite variance and
cause `rem_mv()` to fail. Setting them to `NA` lets the model skip them gracefully (they come back flagged
`error = TRUE` in the result, rather than aborting the whole run).

```{r rem_mv_no_ci, eval=FALSE}
meta_mv <- rem_mv(
  diffexp       = diffexp_list,
  pcriteria     = "pvalue",
  foldchangecol = "log2FC",
  genenamecol   = "Symbol",
  geneidcol     = "Symbol",
  llcol         = NULL,
  rlcol         = NULL,
  vcol          = "var",
  cvar          = FALSE
)
```

### DESeq2

`prepare_deseq2()` maps a `DESeqResults` object straight into MetaVolcanoR's format.

```{r prepare_deseq2_workflow, eval=FALSE}
library(DESeq2)
library(MetaVolcanoR)

dds <- DESeqDataSetFromMatrix(count_matrix, sample_info, design = ~ condition)
dds <- DESeq(dds)

res_treatment1 <- results(dds, contrast = c("condition", "Treatment1", "Control"))
res_treatment2 <- results(dds, contrast = c("condition", "Treatment2", "Control"))
res_treatment3 <- results(dds, contrast = c("condition", "Treatment3", "Control"))

study1 <- prepare_deseq2(res_treatment1)
study2 <- prepare_deseq2(res_treatment2)
study3 <- prepare_deseq2(res_treatment3)

my_studies <- list(
  "Treatment1_vs_Control" = study1,
  "Treatment2_vs_Control" = study2,
  "Treatment3_vs_Control" = study3
)

meta_results <- rem_mv(diffexp = my_studies, metathr = 0.01,
                        outputfolder = tempdir(), draw = "HTML")
meta_results@MetaVolcano
head(meta_results@metaresult)
```

### limma

`prepare_limma()` maps a `limma::topTable()` output straight into MetaVolcanoR's format.

```{r prepare_limma_workflow, eval=FALSE}
library(limma)
library(MetaVolcanoR)

fit <- eBayes(lmFit(expr, design))
limma_toptable <- topTable(fit, coef = 2, number = Inf, confint = TRUE)

study1 <- prepare_limma(limma_toptable)
```

### edgeR

`prepare_edger()` maps an edgeR `topTags()$table` straight into MetaVolcanoR's format.

```{r prepare_edger_workflow, eval=FALSE}
library(edgeR)
library(MetaVolcanoR)

et  <- exactTest(dge)
tt  <- topTags(et, n = Inf)

study1 <- prepare_edger(tt$table)
```

### Swish/fishpond (transcript-level)

If your differential expression analysis was performed at the transcript level using `swish()` from the
fishpond package, `prepare_swish()` maps your results table straight into MetaVolcanoR's format, deriving
an approximate variance from Swish's test statistic — no need to carry the `SummarizedExperiment` object
or its inferential replicates forward.

`swish()` adds `stat`, `log2FC`, `pvalue`, `locfdr`, `qvalue` to `rowData()`. If you extracted your table
with `det <- as.data.frame(mcols(se))`, these columns carry over directly.

```{r prepare_swish, eval=FALSE}
library(fishpond)
library(SummarizedExperiment)

det_1 <- as.data.frame(mcols(se_swish_1))
det_2 <- as.data.frame(mcols(se_swish_2))

# Auto-detects the identifier column (tx_name, tx_id, transcript_id, transcript_name)
study_1 <- prepare_swish(det = det_1)
study_2 <- prepare_swish(det = det_2, tx_col = "tx_name")  # or specify explicitly

my_studies <- list(study1 = study_1, study2 = study_2)

meta_results <- rem_mv(
  diffexp       = my_studies,
  pcriteria     = "pvalue",
  foldchangecol = "Log2FC",
  genenamecol   = "Symbol",
  llcol         = "CI.L",
  rlcol         = "CI.R",
  cvar          = TRUE,
  metathr       = 0.01,
  draw          = "HTML"
)
```

If your identifiers are plain Ensembl transcript IDs, the `Symbol` column will contain values like
`ENST00000456328.2`. The meta-analysis still runs fine — this only affects how labels look on the plot.

### GEO2R

GEO holds thousands of independent studies of comparable conditions, and NCBI's
[GEO2R](https://www.ncbi.nlm.nih.gov/geo/geo2r/) runs a limma differential-expression analysis on any
suitable series directly in the browser, exporting a standardized table (`Gene.symbol`, `logFC`,
`P.Value`, `t`, ...). `prepare_geo2r()` maps that table to MetaVolcanoR's format, deriving the 95% CI from
the moderated t-statistic (`SE = logFC / t`), so a plain GEO2R export can feed the REM without any manual
reformatting:

```{r prepare_geo2r, eval=FALSE}
geo_tab <- read.delim("GSE12345.top.table.tsv", stringsAsFactors = FALSE)
study_1 <- prepare_geo2r(geo_tab)
head(study_1)

# Combine three independent GEO2R tables
diffexplist_geo <- lapply(list(GSE1 = tab1, GSE2 = tab2, GSE3 = tab3), prepare_geo2r)

meta_geo <- rem_mv(
  diffexp       = diffexplist_geo,
  pcriteria     = "pvalue",
  foldchangecol = "Log2FC",
  genenamecol   = "Symbol",
  llcol         = "CI.L",
  rlcol         = "CI.R",
  cvar          = TRUE,
  jobname       = "GEO2R",
  draw          = "PDF"
)
```

## {-}

**Note:** the `prepare_*` functions automatically remove rows with `NA` values, calculate 95% confidence
intervals where possible, format column names to MetaVolcanoR's expected structure, and filter out
infinite values. All of them converge on the same four columns (`Symbol`, `Log2FC`, `pvalue`, and
`CI.L`/`CI.R`), so once your studies are prepared, the rest of the workflow below is identical no matter
which tool they came from.

## Checking your data is correctly formatted

```{r data_check}
data(diffexplist)
head(diffexplist[[1]])
str(diffexplist[[1]])
```

Your prepared data should have these columns: `Symbol` (or another feature identifier), `Log2FC`,
`pvalue`, and — for the REM method — `CI.L`/`CI.R`.

```{r datasets_summary}
cat("Number of datasets:", length(diffexplist), "\n")
cat("Genes per dataset:",
    paste(sapply(diffexplist, nrow), collapse = ", "), "\n")
```

# Implemented meta-analysis approaches

## 1. Random Effects Model (REM)

The REM MetaVolcano summarizes the fold-change of a feature across several studies, taking the per-study
variance into account, and estimates a *summary p-value* — the probability that the summary fold-change is
not different from zero. The `metathr` parameter highlights the top percentage of most consistently
perturbed features, ranked using the [topconfects](https://bioconductor.org/packages/topconfects/)
approach.

Since the current package version, the REM result also reports `randomP.adjust`: `randomP` corrected for
multiple testing across all meta-analyzed features (Benjamini-Hochberg by default; configurable via the
`fdr_method` argument — see `?rem_mv`). Use `randomP.adjust`, not the nominal `randomP`, for feature
selection and for functional enrichment (see below).

```{r rem_basic}
mv_rem <- rem_mv(
  diffexp       = diffexplist,
  pcriteria     = "pvalue",
  foldchangecol = "Log2FC",
  genenamecol   = "Symbol",
  llcol         = "CI.L",
  rlcol         = "CI.R",
  metathr       = 0.01,
  jobname       = "REM_example",
  outputfolder  = tempdir(),
  draw          = "HTML"
)

head(mv_rem@metaresult, 3)
mv_rem@MetaVolcano
```

### Labels and custom colors

When `label_top_n` is used, features are ranked by the topconfects `rank` metric, which balances effect
size and confidence. The `colors` parameter accepts a named vector for the sign-consistency gradient:

```{r rem_labeled, fig.width=11, fig.height=8, out.width="100%"}
mv_rem_labeled <- rem_mv(
  diffexp       = diffexplist,
  pcriteria     = "pvalue",
  foldchangecol = "Log2FC",
  genenamecol   = "Symbol",
  llcol         = "CI.L",
  rlcol         = "CI.R",
  metathr       = 0.01,
  colors        = c(low = "purple", mid = "white", high = "orange", na = "gray80"),
  label_top_n   = 10,
  label_size    = 3.5,
  plot_title    = "REM Meta-Analysis (Top 10)",
  show_legend   = TRUE,
  jobname       = "REM_labeled",
  outputfolder  = tempdir(),
  draw          = "HTML"
)

mv_rem_labeled@MetaVolcano
```

### Forest plot for a specific feature

```{r forest, fig.width=11, fig.height=8, out.width="100%"}
top3 <- mv_rem@metaresult %>%
  filter(!is.na(randomP.adjust)) %>%
  arrange(rank) %>%
  head(3) %>%
  pull(Symbol)

cat("Top 3 features:", paste(top3, collapse = ", "), "\n")

draw_forest(
  remres        = mv_rem,
  gene          = top3[1],
  genecol       = "Symbol",
  foldchangecol = "Log2FC",
  llcol         = "CI.L",
  rlcol         = "CI.R",
  colors        = c(positive = "darkred", negative = "steelblue",
                    neutral = "gray70", reference = "black"),
  point_size    = 3,
  plot_title    = paste(top3[1], "Log2FC across studies"),
  jobname       = "Forest_example",
  outputfolder  = tempdir(),
  draw          = "HTML"
)
```

## 2. Vote-counting approach

The vote-counting MetaVolcano identifies differentially expressed features per study, based on
user-defined *p-value* and *fold-change* thresholds, and displays how many studies call each feature
differentially expressed and in which direction. A feature perturbed in five studies, three of them
downregulated, gets a sign-consistency score of 2 + (-3) = -1.

Labels are selected using `abs(idx)` — the package's own combined metric of frequency and direction
consistency (`idx = ndeg × ddeg`) — rather than `ndeg` alone, which avoids labeling features with
inconsistent direction across studies just because they were tested often:

| Feature | ndeg | ddeg | idx | Rank by `ndeg` alone | Rank by `abs(idx)` |
|---------|------|------|-----|----------------------|---------------------|
| A       | 5    | 5    | 25  | 1st (tied)            | 1st                 |
| B       | 5    | -5   | -25 | 1st (tied)            | 2nd                 |
| C       | 5    | 1    | 5   | 1st (tied)            | 3rd                 |

```{r vote, fig.width=11, fig.height=8, out.width="100%"}
mv_vote <- votecount_mv(
  diffexp       = diffexplist,
  pcriteria     = "pvalue",
  foldchangecol = "Log2FC",
  genenamecol   = "Symbol",
  geneidcol     = NULL,
  pvalue        = 0.05,
  foldchange    = 0,
  metathr       = 0.01,
  label_top_n   = 10,
  label_size    = 3.5,
  plot_title    = "Vote-Counting (Top 10 by |idx|)",
  jobname       = "Vote_example",
  outputfolder  = tempdir(),
  draw          = "HTML"
)

head(mv_vote@metaresult, 3)
mv_vote@MetaVolcano
```

### Feature barplot with custom colors

`draw_featurebar()` accepts a `colors` parameter and, via `mv_vote@featurefreq`, also shows the inverse
cumulative distribution of consistently perturbed features:

```{r featurebar, fig.width=11, fig.height=8, out.width="100%"}
mv_vote@featurefreq
```

## 3. Combining approach (Fisher's method)

The combining MetaVolcano summarizes a feature's fold-change across studies by mean or median, and
combines its per-study p-values with Fisher's method. As with the REM, the result now also reports
`metap.adjust` — `metap` corrected for multiple testing (default Benjamini-Hochberg, configurable via
`fdr_method`) — which is what should drive feature selection rather than the nominal `metap`.

```{r fisher, fig.width=11, fig.height=8, out.width="100%"}
mv_fisher <- combining_mv(
  diffexp       = diffexplist,
  pcriteria     = "pvalue",
  foldchangecol = "Log2FC",
  genenamecol   = "Symbol",
  metafc        = "Mean",
  metathr       = 0.01,
  collaps       = TRUE,
  colors        = c("#083e46", "gray90", "#811820"),
  point_size    = 1.2,
  show_legend   = TRUE,
  plot_title    = "Fisher's Combined Test",
  jobname       = "Fisher_example",
  outputfolder  = tempdir(),
  draw          = "HTML"
)

head(mv_fisher@metaresult, 3)
mv_fisher@MetaVolcano
```

## 4. Functional enrichment — `enrichment_mv()`

`enrichment_mv()` wraps [fGSEA](https://bioconductor.org/packages/fgsea/) and integrates directly with a
REM MetaVolcano result. Gene sets can be supplied manually or downloaded automatically from MSigDB via
[msigdbr](https://cran.r-project.org/package=msigdbr). It automatically ranks by the FDR-adjusted
`randomP.adjust` when present (falling back to `randomP` with a message for REM results produced by an
older package version).

```{r enrich_auto, eval=FALSE}
# GO Biological Process (default category = "C5")
enrich <- enrichment_mv(mv_rem, subcategory = "GO:BP")
enrich$plot
head(enrich$result)
```

Three ranking strategies are available:

| Ranking | Formula | Best for |
|---------|---------|----------|
| `fc` | `randomSummary` | Pathways with large effects |
| `signed_p` | `-log10(p) × sign(FC)` | Pathways with consistent significance |
| `weighted_fc` | `FC × -log10(p)` | Balanced detection |

```{r enrich_rankings, eval=FALSE}
enrich_fc  <- enrichment_mv(mv_rem, ranking = "fc")
enrich_sp  <- enrichment_mv(mv_rem, ranking = "signed_p")
enrich_wfc <- enrichment_mv(mv_rem, ranking = "weighted_fc")
```

```{r enrich_custom, eval=FALSE}
enrich <- enrichment_mv(
  mv_rem,
  category            = "C5",
  subcategory         = "GO:BP",
  ranking             = "weighted_fc",
  plot_padj           = 0.05,
  plot_top_n          = 20,
  clean_pathway_names = TRUE,
  colors              = c(down = "navy", up = "darkred"),
  plot_title          = "GO:BP Enrichment (weighted FC ranking)"
)
enrich$plot
```

Custom gene sets are also supported directly:

```{r enrich_manual, eval=FALSE}
my_pathways <- list(
  "INFLAMMATION" = c("IL6", "TNF", "IL1B", "CXCL8", "CCL2"),
  "ANGIOGENESIS" = c("VEGFA", "FLT1", "KDR", "ANGPT1"),
  "ECM_REMODEL"  = c("MMP2", "MMP9", "MMP12", "COL1A1")
)

enrich_custom <- enrichment_mv(mv_rem, pathways = my_pathways, clean_pathway_names = FALSE)
```

# Additional example: differentially methylated probes

MetaVolcanoR is feature-agnostic: the same three strategies used above for genes apply directly to
Illumina methylation microarray probes, or any other feature type described by an identifier, a log
fold-change, a p-value, and a confidence interval. This section simulates four Illumina 450k datasets
following the format of `limma`'s differential methylation output, to show the full workflow end to end.

```{r meth_setup, eval=requireNamespace("IlluminaHumanMethylation450kanno.ilmn12.hg19", quietly = TRUE)}
library(IlluminaHumanMethylation450kanno.ilmn12.hg19)
library(tibble)
library(tidyr)
library(purrr)
```

For instructive purposes we use a subset of the 450k probe IDs (the first 1000) to generate simulated
per-probe results across 4 datasets:

```{r meth_simulate, eval=requireNamespace("IlluminaHumanMethylation450kanno.ilmn12.hg19", quietly = TRUE)}
ann450k    <- getAnnotation(IlluminaHumanMethylation450kanno.ilmn12.hg19)
ann450k_df <- as.data.frame(ann450k) %>%
  rownames_to_column(var = "Probe_ID")

simulatedResults <- purrr::map_dfr(1:4, ~
  tibble::tibble(
    dataset = paste0("dataset_", .x),
    probes  = ann450k_df$Probe_ID[1:1000]
  ) %>%
  dplyr::mutate(
    logFC     = rnorm(dplyr::n(), mean = -0.01, sd = 1),
    AveExpr   = rnorm(dplyr::n(), mean = -0.75, sd = 2.5),
    t         = rnorm(dplyr::n(), mean = -0.2, sd = 2),
    P.Value   = runif(dplyr::n(), min = 0, max = 1),
    adj.P.Val = p.adjust(P.Value, method = "BH"),
    B         = rnorm(dplyr::n(), mean = -4.00, sd = 3.5),
    deltaBeta = rnorm(dplyr::n(), mean = -0.001, sd = 0.05)
  )
)

head(simulatedResults)
```

MetaVolcanoR needs a standard error and a fold-change confidence interval. Since limma's moderated
t-statistic is `t = logFC / SE`, we recover `SE = logFC / t` and approximate the 95% CI as
`logFC ± 1.96 * SE` — exactly the same relationship used for `prepare_swish()` and `prepare_geo2r()`
above. Results are then split by dataset into a named list, `meta_input`:

```{r meth_prepare, eval=requireNamespace("IlluminaHumanMethylation450kanno.ilmn12.hg19", quietly = TRUE)}
meta_input <- simulatedResults %>%
  dplyr::filter(!is.na(t) & t != 0) %>%
  dplyr::mutate(
    SE   = abs(logFC / t),
    CI.L = logFC - (1.96 * SE),
    CI.R = logFC + (1.96 * SE)
  ) %>%
  dplyr::select(probes, logFC, P.Value, CI.L, CI.R, dataset) %>%
  dplyr::group_split(dataset) %>%
  purrr::set_names(purrr::map_chr(., ~ unique(.x$dataset))) %>%
  purrr::map(as.data.frame)

str(meta_input, max.level = 1)
```

Once formatted, all three MetaVolcanoR strategies apply exactly as with gene-level data:

```{r meth_rem, eval=requireNamespace("IlluminaHumanMethylation450kanno.ilmn12.hg19", quietly = TRUE), fig.width=11, fig.height=8, out.width="100%"}
mv_rem_meth <- rem_mv(
  diffexp       = meta_input,
  pcriteria     = "P.Value",
  foldchangecol = "logFC",
  genenamecol   = "probes",
  llcol         = "CI.L",
  rlcol         = "CI.R",
  metathr       = 0.01,
  colors        = c(low = "#df76ac", mid = "white", high = "#61a110", na = "gray80"),
  label_top_n   = 5,
  plot_title    = "REM — methylation probes",
  jobname       = "REM_methylation",
  outputfolder  = tempdir(),
  draw          = "HTML"
)

mv_rem_meth@MetaVolcano
```

```{r meth_vote, eval=requireNamespace("IlluminaHumanMethylation450kanno.ilmn12.hg19", quietly = TRUE), fig.width=11, fig.height=8, out.width="100%"}
mv_vote_meth <- votecount_mv(
  diffexp       = meta_input,
  pcriteria     = "P.Value",
  foldchangecol = "logFC",
  genenamecol   = "probes",
  geneidcol     = NULL,
  pvalue        = 0.05,
  foldchange    = 0,
  metathr       = 0.01,
  label_top_n   = 10,
  label_size    = 3.5,
  plot_title    = "Vote-Counting — methylation probes",
  jobname       = "Vote_methylation",
  outputfolder  = tempdir(),
  draw          = "HTML"
)

mv_vote_meth@MetaVolcano
```

```{r meth_fisher, eval=requireNamespace("IlluminaHumanMethylation450kanno.ilmn12.hg19", quietly = TRUE), fig.width=11, fig.height=8, out.width="100%"}
mv_fisher_meth <- combining_mv(
  diffexp       = meta_input,
  pcriteria     = "P.Value",
  foldchangecol = "logFC",
  genenamecol   = "probes",
  metafc        = "Mean",
  metathr       = 0.01,
  collaps       = TRUE,
  colors        = c("#df76ac", "gray90", "#61a110"),
  point_size    = 1.2,
  show_legend   = TRUE,
  plot_title    = "Fisher's Combined Test — methylation probes",
  jobname       = "Fisher_methylation",
  outputfolder  = tempdir(),
  draw          = "HTML"
)

mv_fisher_meth@MetaVolcano
```

Since `getAnnotation(IlluminaHumanMethylation450kanno.ilmn12.hg19)` also retrieves probe metadata — which
genomic feature each probe overlaps (promoter, gene body, TSS, UTRs) — it's straightforward to filter for
features of interest before running the meta-analysis, if a specific regulatory context is the target of
the study.

```{r session}
sessionInfo()
```
