---
title: "MetaVolcanoR for proteomics: a pan-cancer meta-analysis of the CPTAC tumor proteome"
author: "Izabela Mamede, Bárbara Marinho, Helder I. Nakaya"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{MetaVolcanoR for proteomics: a pan-cancer meta-analysis of the CPTAC tumor proteome}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(
  tidy    = FALSE,
  warning = FALSE,
  message = FALSE,
  cache   = FALSE,
  fig.width  = 9,
  fig.height = 7,
  out.width  = "100%",
  fig.align  = "center"
)
```

> 📄 See also: [Gene and transcript-level meta-analysis (the main MetaVolcanoR vignette)](GeneExpression.html)

- [Overview](#overview)
    - [CPTAC vignette (optional)](#cptac-vignette-optional)
- [1. Downloading each cohort's proteome](#downloading-each-cohorts-proteome)
- [2. Per-cohort tumor-vs-normal differential abundance](#per-cohort-tumor-vs-normal-differential-abundance)
- [3. Random Effects Model meta-analysis](#random-effects-model-meta-analysis)
    - [Forest plot for a top consensus protein](#forest-plot-for-a-top-consensus-protein)
- [4. Vote-counting and Fisher combining](#vote-counting-and-fisher-combining)
- [5. Functional enrichment of the consensus proteins](#functional-enrichment-of-the-consensus-proteins)
- [Data access and reproducibility](#data-access-and-reproducibility)

## Overview

MetaVolcanoR was first built for gene-expression meta-analysis, but it operates on any table of features
described by an identifier, a log fold change, a p-value, and a variance estimate. This makes it directly
applicable to **differential protein abundance** from mass-spectrometry proteomics.

This vignette demonstrates the complete workflow on **real, harmonized cancer proteomics** from the
Clinical Proteomic Tumor Analysis Consortium (CPTAC) [@cptac; @cptacpkg]. We meta-analyze the
tumor-versus-normal proteome across **three independent cancer cohorts**:

| Label | Cancer | Cohort | Tumor / Normal |
|-------|--------|--------|----------------|
| `CCRCC` | Clear cell renal cell carcinoma | CPTAC-CCRCC | 110 / 84 |
| `LUAD`  | Lung adenocarcinoma             | CPTAC-LUAD  | 111 / 102 |
| `UCEC`  | Uterine corpus endometrial carcinoma | CPTAC-UCEC | 103 / 49 |

Each cohort is an independent set of patients, profiled by the same harmonized TMT pipeline (University of
Michigan / `umich` source), with tumor and normal-adjacent tissue. This is an ideal meta-analysis substrate:
the same biological contrast (tumor vs normal), measured across genuinely different tissues and patient
populations. Proteins that move consistently across all three are **pan-cancer** tumor markers — the robust
signal a meta-analysis is designed to surface, above and beyond any single cancer type. All plots below use
MetaVolcanoR's default styling; see the
[Customizing your plots](GeneExpression.html#customizing-your-plots) section of the main vignette for
colors, titles, and labeling options.

> **A note on method.** CPTAC abundances are log2-ratios to a study-specific reference channel, so raw
> values are **not** comparable across cohorts. We therefore compute the tumor-vs-normal contrast **within
> each cohort first** (with `limma`), which cancels the study-specific reference, and only then
> meta-analyze the resulting effect sizes. This is exactly the situation the Random Effects Model is built
> for.

### CPTAC vignette (optional)

This vignette pulls data from the `cptac` Python package. To build it, create the expected environment
once:

```r
reticulate::virtualenv_create("cptac311", version = "3.11")
reticulate::virtualenv_install("cptac311", packages = "cptac")
```

Without it, the vignette's chunks are skipped and the rest of the site still builds.

```{r packages}
library(reticulate)
library(limma)
library(MetaVolcanoR)
library(dplyr)

# Point reticulate at the Python env where cptac is installed, e.g.:
use_virtualenv("cptac311", required = TRUE)

cptac <- import("cptac")
```

## 1. Downloading each cohort's proteome

Each CPTAC cancer is loaded with its own class, and proteomics comes from the harmonized `umich` source.
Rows are samples, columns are proteins.

```{r load}
loaders <- list(
  CCRCC = cptac$Ccrcc,
  LUAD  = cptac$Luad,
  UCEC  = cptac$Ucec
)

get_prot <- function(loader) {
  ds <- loader()
  as.data.frame(ds$get_proteomics("umich"))
}

proteomes <- lapply(loaders, get_prot)
sapply(proteomes, dim)   # samples x proteins per cohort
```

## 2. Per-cohort tumor-vs-normal differential abundance

In every CPTAC cohort, normal-adjacent samples carry a `.N` suffix on the sample ID, while tumor samples do
not. We use that to define the contrast, then run `limma` (which returns a moderated t-test with confidence
intervals — the `CI.L`/`CI.R` MetaVolcanoR needs for the Random Effects Model).

```{r de-function}
de_tumor_vs_normal <- function(prot) {
  ids   <- rownames(prot)
  group <- ifelse(grepl("\\.N$", ids), "Normal", "Tumor")
  group <- factor(group, levels = c("Normal", "Tumor"))

  # proteins in rows, samples in columns, for limma
  expr <- t(as.matrix(prot))
  # keep proteins quantified in >= 50% of samples
  expr <- expr[rowMeans(!is.na(expr)) >= 0.5, , drop = FALSE]

  design <- model.matrix(~ group)          # coef 2 = Tumor vs Normal
  fit <- eBayes(lmFit(expr, design))
  tt  <- topTable(fit, coef = 2, number = Inf, sort.by = "none",
                  confint = TRUE)

  data.frame(
    Symbol = sub("^\\('([^']+)'.*$", "\\1", rownames(tt)),   # extract gene from ('GENE','ENSP..') tuple
    Log2FC = tt$logFC,
    pvalue = tt$P.Value,
    CI.L   = tt$CI.L,
    CI.R   = tt$CI.R,
    stringsAsFactors = FALSE
  ) |>
    dplyr::filter(is.finite(Log2FC), is.finite(pvalue), nzchar(Symbol)) |>
    dplyr::arrange(pvalue) |>
    dplyr::distinct(Symbol, .keep_all = TRUE)
}

diffprot <- lapply(proteomes, de_tumor_vs_normal)
sapply(diffprot, nrow)         # proteins tested per cohort
lapply(diffprot, head, 3)
```

## 3. Random Effects Model meta-analysis

The REM result reports `randomP.adjust` — the summary p-value corrected for multiple testing across all
meta-analyzed proteins (Benjamini-Hochberg by default; see `?rem_mv` for `fdr_method`). Use it, rather than
the nominal `randomP`, for feature selection and downstream functional enrichment.

```{r rem}
mv_rem <- rem_mv(
  diffexp       = diffprot,
  pcriteria     = "pvalue",
  foldchangecol = "Log2FC",
  genenamecol   = "Symbol",
  llcol         = "CI.L",
  rlcol         = "CI.R",
  metathr       = 0.01,
  label_top_n   = 15,
  label_size    = 2,
  jobname       = "CPTAC_REM",
  outputfolder  = tempdir(),
  draw          = "HTML"
)

mv_rem@MetaVolcano
mv_rem@metaresult |>
  dplyr::arrange(rank) |>
  dplyr::select(Symbol, randomSummary, randomP, randomP.adjust, signcon, rank) |>
  head(20)
```

The top consensus proteins — consistent across renal, lung, and endometrial tumors — are the strongest
pan-cancer candidates. A `signcon` of `3` or `-3` means the protein moved in the same direction in all three
cohorts.

### Forest plot for a top consensus protein

```{r forest}
top_gene <- mv_rem@metaresult$Symbol[116]
draw_forest(
  remres        = mv_rem,
  gene          = top_gene,
  genecol       = "Symbol",
  foldchangecol = "Log2FC",
  llcol         = "CI.L",
  rlcol         = "CI.R",
  jobname       = "CPTAC_forest",
  outputfolder  = tempdir(),
  draw          = "HTML"
)
```

## 4. Vote-counting and Fisher combining

```{r vote}
mv_vote <- votecount_mv(
  diffexp       = diffprot,
  pcriteria     = "pvalue",
  foldchangecol = "Log2FC",
  genenamecol   = "Symbol",
  geneidcol     = NULL,
  pvalue        = 0.05,
  foldchange    = 1,
  metathr       = 0.01,
  label_top_n   = 15,
  jobname       = "CPTAC_vote",
  outputfolder  = tempdir(),
  draw          = "HTML"
)
mv_vote@featurefreq
mv_vote@MetaVolcano
```

```{r fisher}
mv_fisher <- combining_mv(
  diffexp       = diffprot,
  pcriteria     = "pvalue",
  foldchangecol = "Log2FC",
  genenamecol   = "Symbol",
  metafc        = "Mean",
  metathr       = 0.01,
  label_top_n   = 15,
  jobname       = "CPTAC_fisher",
  outputfolder  = tempdir(),
  draw          = "HTML"
)
mv_fisher@MetaVolcano
```

## 5. Functional enrichment of the consensus proteins

```{r enrich, eval=FALSE}
enr <- enrichment_mv(
  mv_rem,
  category    = "H",            # MSigDB Hallmark
  ranking     = "weighted_fc",
  plot_padj   = 0.1,
  plot_top_n  = 20,
  plot_title  = "Hallmark enrichment — pan-cancer proteome meta-analysis"
)
enr$plot
head(enr$result)
```

## Data access and reproducibility

All data are from CPTAC via the open-source `cptac` Python package [@cptacpkg] (`pip install cptac`),
accessed from R through `reticulate`. The proteomics tables are downloaded from the package's public data
repository at run time; no manual downloads are required. MetaVolcanoR is free and open source under GPL-3.

If your network performs SSL inspection and the download fails with a certificate error, enabling the
system certificate store resolves it:

```{r truststore, eval=FALSE}
py_install("truststore", pip = TRUE)
py_run_string("import truststore; truststore.inject_into_ssl()")
```
